---
title: "Job Scheduling R Markdown Reports via R"
output: 
  html_document: 
    highlight: textmate
    toc: yes
    toc_float: 
      collapsed: no
    theme: cosmo
---

## Introduction

Intro to Job Scheduling R Markdown Reports via R.

R is a powerful programming language for statistical computing with many packages and tools.

The goal of this article is to arm you with tools and techniques for job scheduling R Markdown reports via R.

## Use Case

Question: I have a R Markdown report that gathers new data and generates a report. I would rather this be automatic and then output a HTML report to a publishing platform or web server. How can I do this?

## Scheduling

You could create and write your own batch jobs or cron jobs to run your report automagically. Luckly, there are 2 packages that help people do this via R.

http://www.bnosac.be/index.php/blog/64-scheduling-r-scripts-and-processes-on-windows-and-unix-linux

For windows:

https://github.com/bnosac/taskscheduleR

https://cran.r-project.org/web/packages/taskscheduleR/index.html

https://cran.r-project.org/web/packages/taskscheduleR/vignettes/taskscheduleR.html

For Unix/Linux:

https://github.com/bnosac/cronR

https://cran.r-project.org/web/packages/cronR/index.html

https://cran.r-project.org/web/packages/cronR/vignettes/cronR.html

## Example Windows

Install taskscheduleR from cran or github.

1. Create a R script called Jobrun.R with the following text:

```{r}
#library(rmarkdown)
#Sys.setenv(RSTUDIO_PANDOC="C:/Program Files/RStudio/bin/pandoc")
#setwd("C:/Users/Phil/Documents")
#rmarkdown::render("Stocks.Rmd", params = list(symbol='IBM'))
```

pandoc from RSudio Server was not recognized when running script via the Windows task scheduler... So I added Sys.setenv(RSTUDIO_PANDOC="C:/Program Files/RStudio/bin/pandoc") and it worked as above.

https://stackoverflow.com/questions/35624025/pandoc-from-rsudio-server-not-recognized-when-running-script-via-cron

https://stackoverflow.com/questions/28432607/pandoc-version-1-12-3-or-higher-is-required-and-was-not-found-r-shiny

The 'rmarkdown' package will default to using a version of 'pandoc' discovered on the PATH, assuming it meets the pre-requisite version required. If you do not have 'pandoc' on the PATH, you can force RStudio to use your own custom pandoc by setting the environment variable. I did have pandoc and had to add the line for the job to run successfully.

https://support.rstudio.com/hc/en-us/community/posts/239075008-pandoc

If you want, you can put the file in the extdata folder within the package:

R\win-library\3.2\taskscheduleR\extdata

The log file is located at the directory where the R script is located.

## Rendering Output

The job will run a r script that runs the R Markdown report by calling the rmarkdown::render function

http://rmarkdown.rstudio.com/authoring_quick_tour.html#rendering_output

https://www.rdocumentation.org/packages/rmarkdown/versions/1.6/topics/render

## Parameterized Reports

Please note the R Markdown report is Parameterized. This means the YAML Params Field has "IBM"... you could change this to "GOOG" and the same code will generate a new report.

http://rmarkdown.rstudio.com/developer_parameterized_reports.html

https://itsalocke.com/blog/r-quick-tip-parameter-re-use-within-rmarkdown-yaml/

## Output Formats

Please note the R Markdown report is HTML. You could change this to PDF or Word if you like in YAML header.

http://rmarkdown.rstudio.com/lesson-9.html

## knitr Language Engines

Please note if you wanted, you could have your R Markdown file call out to another language when the job runs:

Python, Awk, Ruby, Haskell, Bash, Perl, Graphviz, TikZ, SAS, Scala, and CoffeeScript, etc

http://rmarkdown.rstudio.com/authoring_knitr_engines.html

http://rmarkdown.rstudio.com/authoring_knitr_engines.html

https://yihui.name/knitr/demo/engines/

https://github.com/rstudio/seamonster/tree/master/end-to-end-twitter-etl

Feather can be used to read and write data frames between R and python.

https://github.com/rstudio/seamonster/tree/master/rmarkdown/bash-python-feather

https://cran.r-project.org/web/packages/feather/index.html

You could also connect to the database via R Markdown

https://rviews.rstudio.com/2017/10/18/database-queries-with-r/

{sql, connection=con, output.var = "mydataframe"}
SELECT "month_idx", "year", "month", SUM(CASE WHEN ("term_deposit" = 'yes') THEN (1.0) ELSE (0.0) END) AS "subscribe",
COUNT(*) AS "total"
FROM ("bank") 
GROUP BY "month_idx", "year", "month"

## Generating multiple versions of the R Markdown Report

Above, the function:

rmarkdown::render("Stocks.Rmd", params = list(symbol='IBM'))

Will generate one output file (html)...

What if we wanted multiple reports based on a list of parameters? Report1, Report2, Report3 etc.?

First, lets make a vector of stock symbols:

```{r}
# stocks <- c('IBM', 'GOOG', 'TESLA')
```

Now lets use the purr package map() function to loop through our vector and pass in our stock symbols for 3 reports:

```{r}
# purr::map(stocks, ~rmarkdown::render("Stocks.Rmd", params = list(stocks = .x)))
```

http://purrr.tidyverse.org
https://github.com/tidyverse/purrr
https://cran.r-project.org/web/packages/purrr/index.html

https://www.rstudio.com/resources/webinars/the-tidyverse-and-rstudio-connect/

You can learn more at time 22:49.

2. Create a R Markdown file by going to File -> New File -> R Markdown

Please see Stocks.RMD in this repo as an example.

Put this in the working directory specified above in step 1.

3. Using taskscheduleR, run this code in the console or via the Shiny gadget
*I had issues getting the gadget to work and found running the code directly via the console to work well.

```{r}
# library(taskscheduleR)

# myscript <- system.file("extdata", "Jobrun.R", package = "taskscheduleR")

## run script once within 40 seconds
# taskscheduler_create(taskname = "myfancyscript", rscript = myscript, schedule = "ONCE", starttime = format(Sys.time() + 40, "%H:%M"))
```

This will run the Windows task scheduler once 40 secs after its run.

You can see all jobs with:

```{r}
## get a data.frame of all tasks
# tasks <- taskscheduler_ls()
# str(tasks)
```

And you can delete them with 

```{r}
# taskscheduler_delete(taskname = "myfancyscript")
```

## Example Unix/Linux

https://github.com/bnosac/cronR

Make sure the cron daemon (https://en.wikipedia.org/wiki/Cron) is running. On Debian this is done as follows.

sudo apt-get update
sudo apt-get install -y cron
sudo cron start

Install cronR from cran or github.
*To be updated for RMD soon.
*I had issues getting the gadget to work and found running the code directly via the console to work well.

1. Run this code in the console...

https://cran.r-project.org/web/packages/cronR/cronR.pdf

```{r}
# library(cronR)
# f <- system.file(package = "cronR", "extdata", "helloworld.R")
# cmd <- cron_rscript(f)
# cron_add(command = cmd, frequency = 'minutely', id = 'test1', description = 'My process 1')
# cron_njobs()
# cron_ls()
# cron_rm(id = "test1")
# cron_njobs()
# cron_ls()
```

This will create a cron job to run helloworld.R in the extdata folder.

2. Run this code to check log after run:

https://github.com/bnosac/cronR/blob/master/vignettes/cronR.Rmd


```{r}
## log file is at the place where the helloworld.R script was located
# system.file("extdata", "helloworld.log", package = "cronR")
```

This will run the addin:

https://rdrr.io/cran/cronR/man/cron_rstudioaddin.html

```{r}
# cron_rstudioaddin(RscriptRepository)
```

## Interesting

http://code.markedmondson.me/4-ways-schedule-r-scripts-on-google-cloud-platform/

## Other

https://stackoverflow.com/questions/41819899/scheduled-job-on-cronr-not-running-log-reads-bin-sh-rscript-command-not-fou

https://github.com/rstudio/rmarkdown/issues/620

http://rstudio-pubs-static.s3.amazonaws.com/3269_a6682dfda37e411fb5e0e6699495cdc4.html

https://cran.r-project.org/web/packages/taskscheduleR/README.html

Mark on administrator rights

By default, to schedule a task, you must be a member of the Administrators, Backup Operators, or Server Operators group on the local computer. If you are not, you can ask your System administrator to make sure you have the rights to execute Schtasks.exe. This is the application this R package connects to. Schtasks.exe enables an administrator to create, delete, query, change, run and end scheduled tasks on a local or remote computer. If you are using RStudio, you might need to start RStudio as admin, on Windows this is rightclick >> run as administrator.

https://github.com/rstudio/rmarkdown/issues/319
https://github.com/yihui/knitr/blob/master/inst/bin/knit

https://support.rstudio.com/hc/en-us/community/posts/222207188-Multiple-Python-chunks-in-RStudio-R-Markdown

https://stackoverflow.com/questions/40512587/rstudio-notebook-rmarkdown-running-python-describe-not-shown